Connection between allocation scoring rules an CRPS: example with exponential forecasts

### 2023-04-18 notes/crps_connection_exp.Rmd

---
title: "Connection between allocation scoring rules an CRPS: example with exponential forecasts"
date: "`r Sys.Date()`"
output: 
    pdf_document:
        latex_engine: xelatex
urlcolor: blue
bibliography: allocation.bib 
header-includes:
   - \usepackage{amsmath}
   - \usepackage{algorithm}
   - \usepackage[noend]{algpseudocode}
   - \usepackage{amsthm}
   - \usepackage{bm}
   - \usepackage{cases}
   - \usepackage{caption}
   - \usepackage{unicode-math}
   - \usepackage{hyperref}
   - \DeclareMathOperator*{\argmin}{argmin}
---

@@

If I've correctly translated from the notes document, the expected loss function is
\begin{equation*}
\bar{s}_F(\mathbf{x}) = \sum_{i=1}^2 \mathbb{1}\{0 \leq x_i\}\left[ \sigma e^{-x_i/\sigma} - \sigma \right]
\end{equation*}

I'm going with this below because I think it's close, but I don't actually think it's correct because it leads to negative loss values.

@@

x_1 <- qexp(tau, rate = 1 / sigma_1)
x_2 <- qexp(tau, rate = 1 / sigma_2)

n_grid <- 1001
grid_l <- 0.01
grid_u <- 6.0
y_grid <- tidyr::expand_grid(
    y_1 = seq(from = grid_l, to = grid_u, length.out = n_grid),
    y_2 = seq(from = grid_l, to = grid_u, length.out = n_grid)
)

joint_dist <- y_grid %>%
    dplyr::mutate(
        s_bar = sigma_1 * exp(-y_1 / sigma_1) - sigma_1 +
                sigma_2 * exp(-y_2 / sigma_2) - sigma_2
    )

ggplot(data = joint_dist) +
    geom_raster(aes(x = y_1, y = y_2, fill = s_bar)) +
    geom_contour(mapping = aes(x = y_1, y = y_2, z = s_bar)) +
    geom_abline(intercept = K, slope = -1) +
    geom_point(x = x_1, y = x_2, color = "orange") +
    scale_fill_viridis_c() +
    xlim(0, 6) +
    ylim(0, 6) +
    theme_bw()
```

>>>>

---
title: "Connection between allocation scoring rules an CRPS: example with exponential forecasts"
date: "`r Sys.Date()`"
output: 
    pdf_document:
        latex_engine: xelatex
urlcolor: blue
bibliography: allocation_notes.bib 
header-includes:
   - \usepackage{amsmath}
   - \usepackage{algorithm}
   - \usepackage[noend]{algpseudocode}
   - \usepackage{amsthm}
   - \usepackage{bm}
   - \usepackage{cases}
   - \usepackage{caption}
   - \usepackage{unicode-math}
   - \usepackage{hyperref}
   - \DeclareMathOperator*{\argmin}{argmin}
---

@@

The expected loss function is
\begin{equation*}
\bar{s}_F(\mathbf{x}) = \sum_{i=1}^2 \mathbb{1}\{x_i < 0\}(\sigma - x_i) + \mathbb{1}\{0 \leq x_i\}\sigma e^{-x_i/\sigma}.
\end{equation*}

@@

x_1star <- qexp(tau, rate = 1 / sigma_1)
x_2star <- qexp(tau, rate = 1 / sigma_2)

n_grid <- 101
grid_l <- -2.0
grid_u <- 9.0
y_grid <- tidyr::expand_grid(
    y_1 = seq(from = grid_l, to = grid_u, length.out = n_grid),
    y_2 = seq(from = grid_l, to = grid_u, length.out = n_grid)
)

s_bar <- function(x_1, x_2) {
    sigma_1 * exp(-x_1 / sigma_1)*(x_1>=0) + sigma_2 * exp(-x_2 / sigma_2)*(x_2>=0) +
    (sigma_1 - x_1)*(x_1<0) + (sigma_2 - x_2)*(x_2<0) 
}

joint_dist <- y_grid %>%
    dplyr::mutate(
        s_bar = s_bar(y_1, y_2)
    )

ggplot(data = joint_dist) +
    geom_raster(aes(x = y_1, y = y_2, fill = s_bar)) +
    geom_contour(mapping = aes(x = y_1, y = y_2, z = s_bar), breaks = s_bar(x_1star, x_2star)%%1 + seq(0, 10, by = .5)) +
    geom_abline(intercept = K, slope = -1) +
    geom_point(x = x_1star, y = x_2star, color = "orange") +
    scale_fill_viridis_c() +
    xlim(grid_l, grid_u) +
    ylim(grid_l, grid_u) +
    theme_bw()
```
